Symbolic Execution for Randomized Programs

### 2021-09-04 draft/supplemental.tex
Reviews

\paragraph*{Contributions and outline.}
After illustrating our method on a motivating example (\cref{sec:overview}), we
present our main contributions:
% 
\begin{itemize}
\item A symbolic execution method for probabilistic programs with unknown
  input parameters, along with a formal proof of soundness for the extensions
  to the probabilistic case (\cref{sec:pse}).
\item A broad collection of case studies drawn from the randomized algorithms
  literature (\cref{sec:case_studies}). All case studies have unknown input
  parameters, and none of the examples can be handled using existing automated
  methods. Some example properties include bounding the soundness probability of
  Freivalds' algorithm~\citep{freivalds1977}, a randomized algorithm to check
  matrix multiplication; the expected number of comparisons for randomized
  quicksort; and the correctness probability of a randomized property-testing
  algorithm for checking monotonicity~\citep{goldreich_2017}.
\item An implementation of our approach called \SYSTEM, building on the KLEE
  execution engine~\citep{cadar2008}, and an evaluation on our case studies
  (\cref{sec:impl_and_eval}). By building on KLEE,~\SYSTEM is able to
  perform symbolic execution on real implementations of probabilistic programs
  with natural code written in a wide variety of mainstream languages (e.g.,
  anything targeting LLVM), while faithfully models realistic program states
  (e.g., with overflowing arithmetic, arrays, etc.).
\end{itemize}
% 
We discuss related work in \cref{sec:related}, and future directions in
\cref{sec:conclusion}. \emph{An extended version of the paper that includes all the detailed proofs (in the appendix) is provided as supplemental material.}

@@

We will now prove soundness for \Cref{alg:symb_ex} by showing that our abstraction $R$ respects the semantics defined in the supplemental material for each of the three main types of statements in \textbf{pWhile}, namely \textit{assignment}, \textit{sampling}, and \textit{branching}. A proof for each theorem can be found in the supplemental material.

\begin{restatable}[Correctness of Assignment (\Crefrange{line:beg_raw_assign}{line:end_raw_assign})]{theorem}{soundnessAssignment}
  If a distribution $\mu$ satisfies $R=(\varphi,\sigma,P)$, and~\Cref{alg:symb_ex} reaches an assignment statement of the form $\mathtt{x} \gets e$, then $\mu_{\mathtt{x}\gets e}$, the distribution of program memories after executing $\mathtt{x} \gets e$, as defined in the supplemental material, satisfies $R'=(\varphi,\sigma',P)$, which is produced by~\Cref{alg:symb_ex}.
\end{restatable}

\begin{restatable}[Correctness of Sampling (\Crefrange{line:pse_sym_sample}{line:end_raw_sample})]{theorem}{soundnessSampling}
  If a distribution $\mu$ satisfies $R = (\varphi,\sigma,P)$ and~\Cref{alg:symb_ex} reaches a sampling statement of the form $\mathtt{x} \sim d$, then $\mu_{\mathtt{x} \sim d}$, the distribution of program memories after executing $\mathtt{x} \sim d$, as defined in the supplemental material, satisfies $R'=(\varphi,\sigma',P')$, which is produced by~\Cref{alg:symb_ex}.
\end{restatable}

\begin{restatable}[Correctness of Branching (\Cref{line:pse_sym_branch,line:symbex_true_state,line:symbex_false_state})]{theorem}{soundnessBranching}\label{thm:branch}
  If a distribution $\mu$ satisfies $R=(\varphi,\sigma,P)$, and~\Cref{alg:symb_ex} encounters a branching statement of the form $\mathbf{if}~c~\mathbf{then~goto}~T$, then $\mu_c$ satisfies $R_{true} = (\varphi \wedge c_{sym}, \sigma, P)$ and $\mu_{\neg c}$, the distribution of program memories after executing $\mathbf{if}~c~\mathbf{then~goto}~T$, as defined in the supplemental material, satisfies $R_{false} = (\varphi \wedge \neg c_{sym}, \sigma, P)$. Additionally, for all $a_{\forall} \in \mathit{A}_{\forall}$,
  \[
    \sum_{m \in \{m \in \mathit{M} \mid \deno{c}m = 1\}}\mu(a_{\forall},m) = \deno{p_c} a_{\forall}\quad\text{and}\quad\sum_{m \in \{m \in \mathit{M} \mid \deno{c}m = 0\}}\mu(a_{\forall},m) = \deno{p_c'} a_{\forall}.
  \]
\end{restatable}

@@

\paragraph*{\Q{1} Discussion.}
\begin{table}
  \centering
  \caption{Performance metrics for each of the case studies presented in~\Cref{sec:case_studies}.}
  \label{tab:q1}
  \begin{tabular}{@{}llllllll@{}}
    \toprule
    & \multicolumn{3}{c}{Timing (sec.)} &&\\ \cmidrule{2-4}
    \textbf{Case Study} & \textbf{KLEE} & \textbf{Z3} & \textbf{Total} & \textbf{Lines} & \textbf{Paths} & \textbf{Samples} & \textbf{Concretizations}\\ \midrule
    Freivalds' & 5 & 70 & \textbf{75} & 97 & 8 & 2 & $n=2$ \\
    Freivalds' (Multiple) & 44 & 255 & \textbf{299} & 96 & 8 & 21 & $(n, k) = (3, 7)$ \\
    % Randomized Response (Numer.) & 1.24 & 0.02 & 35 & 3 & 2 & --- \\
    % Randomized Response (Denom.) & 1.21 & 0.02 & 35 & 3 & 2 & --- \\
    Reservoir Sampling & 49 & 79 & \textbf{128} & 52 & 63 & 5 & $(n, k) = (10, 5)$ \\
    Monotone Testing & 4 & 352 & \textbf{356} & 69 & 30 & 1 & $n=23$ \\
    QuickSort & 95 & 489 & \textbf{584} & 65 & 120 & 10 & $n=5$ \\
    Bloom Filter & 256 & 231 & \textbf{487} & 386 & 83 & 8 & $(m, \epsilon) = (3, 0.4)$ \\
    Count-min Sketch & 108 & 206 & \textbf{314} & 245 & 2 & 8 & $(n, \epsilon, \gamma) = (4, 0.5, 0.25)$ \\

@@

\paragraph*{\Q{2} Discussion.}
\begin{table}
  \centering
  \caption{Performance metrics of four different ways of specifying $A \times B \neq C$ in the query for Freivalds' Algorithm (\Cref{sec:freivalds}). Here, $n=2$ and all elements of the matrices are C++ \texttt{int}s.}
  \label{tab:q2}
  \begin{tabular}{@{}lllll@{}}
    \toprule
    & \multicolumn{3}{c}{Timing (sec.)} &\\ \cmidrule{2-4}
    \textbf{Spec. for $A \times B \neq C$} & \textbf{KLEE} & \textbf{Z3} & \textbf{Total} & \textbf{Paths}\\ \midrule
    \textsc{AllOff} & 2 & 1 & \textbf{3} & 2\\
    \textsc{SomeOff}& 5 & 70 & \textbf{75} & 8\\
    \textsc{FirstOff}& 1 & 5 & \textbf{6} & 2\\
    \textsc{OneOff} & 2 & 17 & \textbf{19} & 2\\
    \bottomrule
  \end{tabular}
\end{table}
To better understand~\SYSTEM's performance we consider how changes to the phrasing of the query for Freivalds' Algorithm affects performance.
% 
Recall that Freivalds' Algorithm efficiently determines whether $A \times B = C$, where $A,B$, and $C$ are $n \times n$ matrices; however, it has a false-positive error rate of at most $\frac{1}{2}$.
% 
In~\Cref{eq:freivalds} we set $\psi := A \times B \neq C \wedge \textsc{Freivalds}(A,B,C,n) = \mathtt{true}$ but did not state exactly how $A \times B$ must differ from $C$.

@@

%% Bibliography
\bibliography{header,main}
\end{document}

>>>>

\paragraph*{Contributions and outline.}
After illustrating our method on a motivating example (\cref{sec:overview}), we
present our main contributions:
% 
\begin{itemize}
\item A symbolic execution method for probabilistic programs with unknown
  input parameters, along with a formal proof of soundness for the extensions
  to the probabilistic case (\cref{sec:pse}).
\item A broad collection of case studies drawn from the randomized algorithms
  literature (\cref{sec:case_studies}). All case studies have unknown input
  parameters, and none of the examples can be handled using existing automated
  methods. Some example properties include bounding the soundness probability of
  Freivalds' algorithm~\citep{freivalds1977}, a randomized algorithm to check
  matrix multiplication; the expected number of comparisons for randomized
  quicksort; and the correctness probability of a randomized property-testing
  algorithm for checking monotonicity~\citep{goldreich_2017}.
\item An implementation of our approach called \SYSTEM, building on the KLEE
  execution engine~\citep{cadar2008}, and an evaluation on our case studies
  (\cref{sec:impl_and_eval}). By building on KLEE,~\SYSTEM is able to
  perform symbolic execution on real implementations of probabilistic programs
  with natural code written in a wide variety of mainstream languages (e.g.,
  anything targeting LLVM), while faithfully models realistic program states
  (e.g., with overflowing arithmetic, arrays, etc.).
\end{itemize}
% 
We discuss related work in \cref{sec:related}, and future directions in
\cref{sec:conclusion}.

@@

We will now prove soundness for \Cref{alg:symb_ex} by showing that our abstraction $R$ respects the semantics defined in \Cref{sec:semantics} for each of the three main types of statements in \textbf{pWhile}, namely \textit{assignment}, \textit{sampling}, and \textit{branching}. A proof for each theorem can be found in~\Cref{sec:omitted_proofs}.

\begin{restatable}[Correctness of Assignment (\Crefrange{line:beg_raw_assign}{line:end_raw_assign})]{theorem}{soundnessAssignment}
  If a distribution $\mu$ satisfies $R=(\varphi,\sigma,P)$, and~\Cref{alg:symb_ex} reaches an assignment statement of the form $\mathtt{x} \gets e$, then $\mu_{\mathtt{x}\gets e}$, the distribution of program memories after executing $\mathtt{x} \gets e$, as defined in~\Cref{sec:semantics},  satisfies $R'=(\varphi,\sigma',P)$, which is produced by~\Cref{alg:symb_ex}.
\end{restatable}

\begin{restatable}[Correctness of Sampling (\Crefrange{line:pse_sym_sample}{line:end_raw_sample})]{theorem}{soundnessSampling}
  If a distribution $\mu$ satisfies $R = (\varphi,\sigma,P)$ and~\Cref{alg:symb_ex} reaches a sampling statement of the form $\mathtt{x} \sim d$, then $\mu_{\mathtt{x} \sim d}$, the distribution of program memories after executing $\mathtt{x} \sim d$, as defined in~\Cref{sec:semantics}, satisfies $R'=(\varphi,\sigma',P')$, which is produced by~\Cref{alg:symb_ex}.
\end{restatable}

\begin{restatable}[Correctness of Branching (\Cref{line:pse_sym_branch,line:symbex_true_state,line:symbex_false_state})]{theorem}{soundnessBranching}\label{thm:branch}
  If a distribution $\mu$ satisfies $R=(\varphi,\sigma,P)$, and~\Cref{alg:symb_ex} encounters a branching statement of the form $\mathbf{if}~c~\mathbf{then~goto}~T$, then $\mu_c$ satisfies $R_{true} = (\varphi \wedge c_{sym}, \sigma, P)$ and $\mu_{\neg c}$, the distribution of program memories after executing $\mathbf{if}~c~\mathbf{then~goto}~T$, as defined in~\Cref{sec:semantics}, satisfies $R_{false} = (\varphi \wedge \neg c_{sym}, \sigma, P)$. Additionally, for all $a_{\forall} \in \mathit{A}_{\forall}$,
  \[
    \sum_{m \in \{m \in \mathit{M} \mid \deno{c}m = 1\}}\mu(a_{\forall},m) = \deno{p_c} a_{\forall}\quad\text{and}\quad\sum_{m \in \{m \in \mathit{M} \mid \deno{c}m = 0\}}\mu(a_{\forall},m) = \deno{p_c'} a_{\forall}.
  \]
\end{restatable}

@@

\paragraph*{\Q{1} Discussion.}
% \label{sec:eval_q1}

\begin{table}
  \centering
  \caption{Performance metrics for each of the case studies presented in~\Cref{sec:case_studies}.}
  \label{tab:q1}
  \begin{tabular}{@{}llllllll@{}}
    \toprule
    & \multicolumn{3}{c}{Timing (sec.)} &&\\ \cmidrule{2-4}
    \textbf{Case Study} & \textbf{KLEE} & \textbf{Z3} & \textbf{Total} & \textbf{Lines} & \textbf{Paths} & \textbf{Samples} & \textbf{Concretizations}\\ \midrule
    Freivalds' & 5 & 70 & \textbf{75} & 97 & 8 & 2 & $n=2$ \\
    Freivalds' (Multiple) & 44 & 255 & \textbf{299} & 96 & 8 & 21 & $(n, k) = (3, 7)$ \\
    % Randomized Response (Numer.) & 1.24 & 0.02 & 35 & 3 & 2 & --- \\
    % Randomized Response (Denom.) & 1.21 & 0.02 & 35 & 3 & 2 & --- \\
    Reservoir Sampling & 49 & 79 & \textbf{128} & 52 & 63 & 5 & $(n, k) = (10, 5)$ \\
    Monotone Testing & 4 & 352 & \textbf{356} & 69 & 30 & 1 & $n=23$ \\
    QuickSort & 95 & 489 & \textbf{584} & 65 & 120 & 10 & $n=5$ \\
    Bloom Filter & 256 & 231 & \textbf{487} & 386 & 83 & 8 & $(m, \epsilon) = (3, 0.4)$ \\
    Count-min Sketch & 108 & 206 & \textbf{314} & 245 & 2 & 8 & $(n, \epsilon, \gamma) = (4, 0.5, 0.25)$ \\

@@

\paragraph*{\Q{2} Discussion.}
% \label{sec:eval_q2}

\begin{table}
  \centering
  \caption{Performance metrics of four different ways of specifying $A \times B \neq C$ in the query for Freivalds' Algorithm (\Cref{sec:freivalds}). Here, $n=2$ and all elements of the matrices are C++ \texttt{int}s.}
  \label{tab:q2}
  \begin{tabular}{@{}lllll@{}}
    \toprule
    & \multicolumn{3}{c}{Timing (sec.)} &\\ \cmidrule{2-4}
    \textbf{Spec. for $A \times B \neq C$} & \textbf{KLEE} & \textbf{Z3} & \textbf{Total} & \textbf{Paths}\\ \midrule
    \textsc{AllOff} & 2 & 1 & \textbf{3} & 2\\
    \textsc{SomeOff}& 5 & 70 & \textbf{75} & 8\\
    \textsc{FirstOff}& 1 & 5 & \textbf{6} & 2\\
    \textsc{OneOff} & 2 & 17 & \textbf{19} & 2\\
    \bottomrule
  \end{tabular}
\end{table}
To better understand~\SYSTEM's performance we consider how changes to the phrasing of the query for Freivalds' Algorithm affects performance.
% 
Recall that Freivalds' Algorithm efficiently determines whether $A \times B = C$, where $A,B$, and $C$ are $n \times n$ matrices; however, it has a false-positive error rate of at most $\frac{1}{2}$.
% 
In~\Cref{eq:freivalds} we set $\psi := A \times B \neq C \wedge \textsc{Freivalds}(A,B,C,n) = \mathtt{true}$ but did not state exactly how $A \times B$ must differ from $C$.

@@

%% Bibliography
\bibliography{header,main}

%% Appendix
\appendix
\allowdisplaybreaks

\section{Program Semantics}
\label{sec:semantics}
Here we define the semantics for the three main types of statements which concerns probabilistic symbolic execution: assignments, probabilistic sampling, and branching.

\begin{definition}[Assignment Semantics]
  \label{def:assignment}
  Let $\mu \in \mathit{Dists_M}$ be a distribution of program memories parameterized by assignments to universal symbolic variables. Let $\mathtt{x} \gets e$ be an arbitrary assignment statement where $\mathtt{x} \in \mathcal{X}$ is assigned to the program expression $e$. Let $\mathsf{assign}_{\mathtt{x}\gets e} : \mathit{M} \rightarrow \mathit{M}$ is defined as
  \[
    \mathsf{assign}_{\mathtt{x}\gets e}(m) = \lambda(\mathtt{y} : \mathcal{X})
    \begin{cases}
      \deno{e}m     & \text{if $\mathtt{x} = \mathtt{y}$} \\
      m(\mathtt{y}) & \text{otherwise}
    \end{cases}
  \]
  and let $\mathsf{unassign}_{\mathtt{x}\gets e} = \mathsf{assign}_{\mathtt{x}\gets e}^{-1}$. Define $\mu_{\mathtt{x}\gets e}$ to be the distribution over program memories parameterized by assignments to universal symbolic variables after executing the assignment statement $\mathtt{x}\gets e$ to be
  \[
    \mu_{\mathtt{x}\gets e}(a_{\forall},m) = \sum_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \mu(a_{\forall},m').
  \]
\end{definition}

\begin{definition}[Sampling Semantics]
  \label{def:sampling}
  Let $\mu \in \mathit{Dists_M}$ be a distribution of program memories parameterized by assignments to universal symbolic variables. Let $\mathtt{x} \sim d$ be an arbitrary sampling instruction which assigns the program variable $\mathtt{x} \in \mathcal{X}$ to a random element from the distribution of values parameterized by a memory, represented as a distribution expression $d \in \mathit{Dists}$. Let $\mathsf{desample} : \mathcal{X} \times \mathit{M} \rightarrow \mathcal{P}(\mathit{M})$ be defined as
  \[
    \mathsf{desample}(\mathtt{x},m) = \{ m'\in \mathit{M} \mid \forall (\mathtt{y} \in \mathcal{X})~.~(\mathtt{y} \neq \mathtt{x} \wedge m'(\mathtt{y}) = m(\mathtt{y}))\}.
  \]
  Define $\mu_{\mathtt{x} \sim d}$ to be the distribution over program memories parameterized by assignments to universal symbolic variables after executing the sampling statement $\mathtt{x} \sim d$ to be
  \[
    \mu_{\mathtt{x} \sim d}(a_{\forall},m) = \sum_{m' \in \mathsf{desample}(m)} (\deno{d}a_{\forall})(m(\mathtt{x})) \cdot \mu(a_{\forall},m').
  \]
\end{definition}

\begin{definition}[Branching Semantics]
  \label{def:branching}
  Let $a_{\forall} \in \mathit{A}_{\forall}$ be an arbitrary assignment of universal symbolic variables, $c$ be a guard of an \textbf{if-goto} statement, $\mu \in \mathit{Dists_M}$ be a distribution of program memories parameterized by assignments to universal symbolic variables, and $\mathtt{x}_1,\ldots,\mathtt{x}_n \in \mathcal{X}$ be all the program variables in $c$. Then for all program memories $m \in \mathit{M}$, $\mu$ conditioned on a guard $c$ being true, represented as $\mu_c$ is defined as
  \[
    \mu_c(a_{\forall},m) = \frac{\displaystyle\Pr_{m' \sim \mu(a_{\forall})}\left[ m' = m \land \deno{c}m' = \mathtt{true} \right]}{\displaystyle\Pr_{m' \sim \mu(a_{\forall})}\left[ \deno{c}m' = \mathtt{true} \right]}
  \]
  Similarly, for all program memories $m \in \mathit{M}$, $\mu$ conditioned on a guard $c$ being false, represented as $\mu_{\neg c}$ is defined as
  \[
    \mu_{\neg c}(a_{\forall},m) = \frac{\displaystyle\Pr_{m' \sim \mu(a_{\forall})}\left[ m' = m \land \deno{c}m' = \mathtt{false} \right]}{\displaystyle\Pr_{m' \sim \mu(a_{\forall})}\left[ \deno{c}m' = \mathtt{false} \right]}
  \]
\end{definition}

\section{Omitted Proofs}
\label{sec:omitted_proofs}
\equivpthm*
\begin{proof}
  Let $S$ be an arbitrary \textbf{pWhile} program statement, $\varphi$ be the current path condition, $P$ be the current distribution map, and $p$ be the current path probability before executing $S$.
  % 
  Assume that
  \[
    p = \frac{\displaystyle\sum_{v_1,\ldots,v_n \in \mathcal{D}} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_n \mapsto v_n\} ]}{\abs{\mathcal{D}}}.
  \]
  We will proceed with a case analysis on $S$:
  \begin{itemize}
  \item \textit{Assignment (\Crefrange{line:beg_raw_assign}{line:end_raw_assign})}. Suppose $S$ is an assignment statement of the form $\mathtt{x} \gets e$, where \texttt{x} is an arbitrary program variable and $e$ is an arbitrary program expression.
    % 
    Note that $\varphi$, $P$, and $p$ do not change after executing $\mathtt{x} \gets e$ and so the statement trivially holds.
  \item \textit{Sampling (\Crefrange{line:pse_sym_sample}{line:end_raw_sample})}. Suppose $S$ is a sampling statement of the form $\mathtt{x} \sim d$, where \texttt{x} is an arbitrary program variable and $d$ is an arbitrary distribution expression.
    % 
    Additionally assume that there are $n$ probabilistic symbolic variables in $P$, namely $\delta_1,\ldots,\delta_n$, and $\delta_{n+1}$ is the fresh probabilistic symbolic variable created on \cref{line:fresh_psv} in \Cref{alg:sample}.
    % 
    Then, $P_1 = P \cup \{\delta_{n+1} \mapsto d\}$, according to \cref{line:update_P} where $P_1$ is the updated distribution map returned by \textbf{PSESample} on \cref{line:pse_sym_sample}.
    % 
    If $\mathcal{D}_1 = \dom{P_1[\delta_1]} \times \cdots \times \dom{P_1[\delta_{n+1}]}$, then
    \begin{align*}
      \frac{\displaystyle\sum_{v_1,\ldots,v_{n+1} \in \mathcal{D}_1} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_{n+1} \mapsto v_{n+1}\}]}{\abs{\mathcal{D}_1}} &= \frac{\abs{\dom{P_1[\delta_{n+1}]}} \cdot \displaystyle\sum_{v_1,\ldots,v_{n} \in \mathcal{D}} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_{n} \mapsto v_{n}\}]}{\abs{\mathcal{D}_1}} \\
                                                                                                                         &= \frac{\abs{\dom{P_1[\delta_{n+1}]}} \cdot \displaystyle\sum_{v_1,\ldots,v_{n} \in \mathcal{D}} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_{n} \mapsto v_{n}\}]}{\abs{\dom{P_1[\delta_{n+1}]}} \cdot \abs{\mathcal{D}}} \\
                                                                                                                         &= \frac{\displaystyle\sum_{v_1,\ldots,v_{n} \in \mathcal{D}} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_{n} \mapsto v_{n}\}]}{\abs{\mathcal{D}}} \\
                                                                                                                         &= p
    \end{align*}
    where the first line follows from the fact that $\delta_{n+1}$ cannot appear in $\varphi$ as $\delta_{n+1}$ is a \textit{fresh} probabilistic symbolic variable.
  \item \textit{Branching (\Cref{line:pse_sym_branch,line:symbex_true_state,line:symbex_false_state})}. Suppose $S$ is a branching statement of the form $\mathbf{if}~c~\mathbf{then~goto}~T$ where $c$ is an arbitrary program guard.
    % 
    Additionally assume that there are $n$ probabilistic symbolic variables in $P$, namely $\delta_1,\ldots,\delta_n$.
    % 
    Then, according to \cref{line:branch_prob_compute} of \Cref{alg:branch},
    \[
      p_c = \frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}} [(c_{sym} \wedge \varphi)\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}}[\varphi\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}
    \]
    and so $p_t = p_c$ and $p_f = 1-p_c$.
    % 
    If $p'_t = p \cdot p_t$ and $p'_f = p \cdot p_f$, as on lines \cref{line:symbex_true_state,line:symbex_false_state}, and $\varphi'_t = \varphi \wedge c_{sym}$ and $\varphi'_f = \varphi \wedge \neg c_{sym}$, then we claim that
    \[
      p'_t = \frac{\displaystyle\sum_{v_1,\ldots,v_n \in \mathcal{D}} [ \varphi'_t\{\delta_1 \mapsto v_1, \ldots, \delta_n \mapsto v_n\} ]}{\abs{\mathcal{D}}}
    \]
    and similarly
    \[
      p'_f = \frac{\displaystyle\sum_{v_1,\ldots,v_n \in \mathcal{D}} [ \varphi'_f\{\delta_1 \mapsto v_1, \ldots, \delta_n \mapsto v_n\} ]}{\abs{\mathcal{D}}}.
    \]

    We will prove the claim for $p'_t$ as the proof for $p'_f$ is symmetric.
    % 
    Note that,
    \begin{align*}
      p'_t &= p \cdot p_t \\
           &= \frac{\displaystyle\sum_{v_1,\ldots,v_n \in \mathcal{D}} [ \varphi\{\delta_1 \mapsto v_1, \ldots, \delta_n \mapsto v_n\} ]}{\abs{\mathcal{D}}} \cdot \frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}} [(c_{sym} \wedge \varphi)\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}}[\varphi\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]} \\
           &= \frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}} [(c_{sym} \wedge \varphi)\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\abs{\mathcal{D}}}\\
           &= \frac{\displaystyle\sum_{(v_1,\ldots,v_n) \in \mathcal{D}} [\varphi'_t\{\delta_1 \mapsto v_1,\ldots,\delta_n \mapsto v_n\}]}{\abs{\mathcal{D}}}
    \end{align*}
  \end{itemize}
  Since no state changes if $S$ is a \textbf{halt} statement, the theorem is trivially true for this case.
\end{proof}

\soundnessAssignment*
\begin{proof}
  First assume that $\mu$ satisfies the input abstraction $R=(\varphi,\sigma,P)$. Note that by~\Cref{def:assignment},
  \[
    \mu_{\mathtt{x}\gets e}(a_{\forall},m) = \displaystyle\sum_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \mu(a_{\forall},m')
  \]
  for all $a_{\forall} \in \mathit{A}_{\forall}$ and $m \in \mathit{M}$. Also note that by~\Cref{def:main},
  \[
    \mu(a_{\forall},m) = \displaystyle\sum_{a_p \in \mathsf{fromMem}(\sigma,a_{\forall},m)} \nu_{\varphi}(a_{\forall},a_p).
  \]
  So,
  \begin{align*}
    \mu_{\mathtt{x}\gets e}(a_{\forall},m) &= \sum_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \mu(a_{\forall},m')\\
                                           &= \sum_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \sum_{a_p \in \mathsf{fromMem}(\sigma,a_{\forall},m')} \nu_{\varphi}(a_{\forall},a_p).
  \end{align*}

  Thus, in order to satisfy~\Cref{def:main} it suffices to show that
  \[
    \sum_{a_p \in \mathsf{fromMem}(\sigma',a_{\forall},m)} \nu_{\varphi}(a_{\forall},a_p) = \sum_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \sum_{a_p \in \mathsf{fromMem}(\sigma,a_{\forall},m')} \nu_{\varphi}(a_{\forall},a_p).
  \]

  Note that if $m_1,m_2 \in \mathit{M}$ where $m_1 \neq m_2$, then for all $a_{\forall} \in \mathit{A}_{\forall}$ and $\sigma$, $\mathsf{fromMem}(\sigma,a_{\forall},m_1)$ and $\mathsf{fromMem}(\sigma,a_{\forall},m_2)$ are disjoint. Hence, it suffices to show that
  \[
    \mathsf{fromMem}(\sigma',a_{\forall},m) = \bigcup_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \mathsf{fromMem}(\sigma,a_{\forall},m').
  \]
  to prove the theorem. We will show this through double containment.

  First, let $a_p \in \bigcup_{m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)} \mathsf{fromMem}(\sigma,a_{\forall},m')$ be arbitrary. This means that for some $m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)$, $m' = \lambda (\mathtt{y} : \mathcal{X}) . \deno{\sigma(\mathtt{y})}a_{\forall}a_p$. Furthermore, since $m' \in \mathsf{unassign}_{\mathtt{x}\gets e}(m)$,
  \[
    m = \lambda (\mathtt{y} : \mathcal{X})
    \begin{cases}
      \deno{e}m' & \text{if $\mathtt{x}=\mathtt{y}$}\\
      m'(\mathtt{y}) & \text{otherwise}
    \end{cases}
  \]
  Note that we need to show that $a_p \in \mathsf{fromMem}(\sigma',a_{\forall},m)$. This is equivalent to showing that $m =\mathsf{toMem}(\sigma',a_{\forall},a_p)$. Thus, it is sufficient to prove that
  \[
    \lambda (\mathtt{y} : \mathcal{X}) . \deno{\sigma'(\mathtt{y})}a_{\forall}a_p = \lambda (\mathtt{y} : \mathcal{X})
    \begin{cases}
      \deno{e}m' & \text{if $\mathtt{x}=\mathtt{y}$}\\
      m'(\mathtt{y}) & \text{otherwise}
    \end{cases}.
  \]
  Note that the domains of each of these functions are equal, namely the set of all program variables present within the program. Now we need to show that for every $\mathtt{z} \in \mathcal{X}$, the two functions compute the same result. We have two cases:
  \begin{enumerate}
  \item \textbf{Case 1:} Assume that $\mathtt{z} = \mathtt{x}$. Then we have to show that $\deno{e}m' = \deno{\sigma'(\mathtt{x})}a_{\forall}a_p$. However, note that by~\Crefrange{line:beg_raw_assign}{line:end_raw_assign}, $\sigma'(\mathtt{x}) = e_{sym}$. Recall that $m' = \lambda (\mathtt{y} : \mathcal{X}) . \deno{\sigma(x)}a_{\forall}a_p$. If $\mathtt{x}_1,\ldots,\mathtt{x}_k \in \mathcal{X}$ are all the program variables present, then
    \begin{align*}
      \deno{e}m' &= \mathsf{eval}(e[\mathtt{x}_1 \mapsto \deno{\sigma(\mathtt{x}_1)}a_{\forall}a_p, \ldots, \mathtt{x}_k \mapsto \deno{\sigma(\mathtt{x}_k)}a_{\forall}a_p])
    \end{align*}
    However, note that $e_{sym} = e[\mathtt{x}_1 \mapsto \sigma(\mathtt{x}_1),\ldots,\mathtt{x}_k \mapsto \sigma(\mathtt{x}_k)]$ and so,
    \begin{align*}
      \deno{e_{sym}}a_{\forall}a_p &= \deno{e[\mathtt{x}_1 \mapsto \sigma(\mathtt{x}_1),\ldots,\mathtt{x}_k \mapsto \sigma(\mathtt{x}_k)]}a_{\forall}a_p\\
                             &= \mathsf{eval}(e[\mathtt{x}_1 \mapsto \deno{\sigma(\mathtt{x}_1)}a_{\forall}a_p, \ldots, \mathtt{x}_k \mapsto \deno{\sigma(\mathtt{x}_k)}a_{\forall}a_p]).
    \end{align*}
    Hence, $\deno{e}m' = \deno{\sigma'(\mathtt{x})}a_{\forall}a_p$.
  \item \textbf{Case 2:} Now assume that $\mathtt{z} \neq \mathtt{x}$. We claim that $m'(\mathtt{z}) = \deno{\sigma'(\mathtt{z})}a_{\forall}a_p$. Note that $m'(\mathtt{z}) = \deno{\sigma(\mathtt{z})}a_{\forall}a_p$. Since $\mathtt{z}\neq\mathtt{x}$, then $\sigma(\mathtt{z}) = \sigma'(\mathtt{z})$ and we are done.
  \end{enumerate}
  Therefore,
  \[
    \lambda (\mathtt{y} : \mathcal{X}) . \deno{\sigma'(\mathtt{y})}a_{\forall}a_p = \lambda (\mathtt{y} : \mathcal{X})
    \begin{cases}
      \deno{e}m' & \text{if $\mathtt{x}=\mathtt{y}$}\\
      m'(\mathtt{y}) & \text{otherwise}
    \end{cases}
  \] and so $a_p \in \mathsf{fromMem}(\sigma',a_{\forall},m)$.
